---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Load/install required libraries. Here, we will use SpatialPCA, a dimension reduction method that uses spatial correlation information to extract low-dimensional representations of ST data, facilitating effective clustering and visualization of spatial domains.

We will also try Banksy. It uses a spatial feature augmentation strategy to improve clustering performance by embedding cells in a space that considers both their own and their local neighborhood transcriptomes.

A few alternative methods implemented in R to consider:

BayesSpace: A statistical model that performs spatial clustering using a t-distributed error model and Markov chain Monte Carlo (MCMC) for parameter estimation. It is designed to identify clusters at the spot level, focusing on high-resolution clustering within tissue sections. https://www.bioconductor.org/packages/release/bioc/html/BayesSpace.html

BASS (Bayesian Analysis for Spatial Transcriptomics): This method simultaneously detects spatial domains and clusters cell types within tissue sections using a hierarchical Bayesian model. It is particularly effective for multi-slice clustering, providing robust performance in identifying spatial domains. https://github.com/zhengli09/BASS

Install packages if missing:
```{r}
remotes::install_github("prabhakarlab/Banksy")
remotes::install_github("shangll123/SpatialPCA")

```


```{r}
library(SpatialPCA)
library(Seurat)
library(Banksy)
library(SeuratWrappers)
```

Lets read in some data - we will use a mouse intestine visium Seurat object from earlier:
```{r}
seurat <- readRDS("/project/shared/spatial_data_camp/notebooks_and_code/EXTRA/R_spatial_clustering/seurat.RDS")
```

Let's quickly visualise the clustering solution we had identified via Seurat:
```{r}
SpatialDimPlot(seurat, group.by = "seurat_clusters")

```
To calculate spatial PCA, we first need to construct a spatial PCA object. 

counts=GetAssayData(seurat, assay="Spatial", layer="counts"):

This extracts the count data from a Seurat object (seurat).
GetAssayData() is used to access the data within a specific assay in Seurat (in this case, "Spatial"), and layer="counts" specifies that the raw count data should be retrieved. This data serves as the input for creating the spatial PCA object.


location=as.matrix(GetTissueCoordinates(seurat)[, 1:2]):

This specifies the spatial coordinates of each spot or cell in the tissue.
GetTissueCoordinates(seurat) retrieves the tissue coordinates from the Seurat object, and [, 1:2] selects the first two columns,representing the x and y coordinates in the tissue. as.matrix() converts these coordinates into a matrix format, which is required by the CreateSpatialPCAObject function.


```{r}
ST <- CreateSpatialPCAObject(counts=GetAssayData(seurat, assay="Spatial", layer="counts"),
                            location=as.matrix(GetTissueCoordinates(seurat)[, 1:2]),
                            project = "SpatialPCA",
                            gene.type="spatial",
                            customGenelist=NULL,
                            min.loctions = 20,
                            min.features=20)
```


```{r}
ST <- SpatialPCA_buildKernel(ST)

ST <- SpatialPCA_EstimateLoading(ST,
                                 fast=TRUE,
                                 SpatialPCnum=50)

ST <- SpatialPCA_SpatialPCs(ST, 
                            fast=TRUE)
```

```{r}
seurat[["spatialpca"]] <- CreateDimReducObject(embeddings=t(ST@SpatialPCs), 
                                               key = "spatialpca",
                                               assay="Spatial")
```

```{r}
seurat <- FindNeighbors(seurat, reduction = "spatialpca", dims = 1:20)
seurat <- FindClusters(seurat, resolution = .7, cluster.name = "spatialpca_clusters")
```

```{r}
SpatialDimPlot(seurat, group.by = "spatialpca_clusters")
```

Next, we can try calculating clusters using Banksy
```{r}
seurat <- RunBanksy(seurat,
                    lambda = 0.2,
                    verbose=TRUE, 
                    assay = 'Spatial',
                    slot = 'data',
                    features = 'variable',
                    k_geom = 15)

seurat <- RunPCA(seurat, assay = 'BANKSY', features = rownames(seurat))
seurat <- RunUMAP(seurat, dims = 1:20)
seurat <- FindNeighbors(seurat, dims = 1:20)
seurat <- FindClusters(seurat, resolution = 0.7, cluster.name="banksy_clusters")
```

```{r}
SpatialDimPlot(seurat, group.by = "banksy_clusters")
```